---
title: "data_manipulation with "dplyr""
output: github_document
date: "2022-09-22"
---
library(tidyverse)
```{r setup, include=FALSE}
options(tibble.print_min = 3)

litters_data = read_csv("data_import_examples/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("data_import_examples/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

## "select"

#You can specify the columns you want to keep by naming all of them:
```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

#You can specify the specify a range of columns to keep:
```{r}
select(litters_data, group:gd_of_birth)
```

#You can also specify columns you’d like to remove:
```{r}
select(litters_data, -pups_survive)
```

#:You can rename variables as part of this process:
```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

#If all you want to do is rename something, you can use rename instead of select. 
```{r}
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```
# select: starts_with(), ends_with(), and contains()

```{r}
select(litters_data, starts_with("gd"))
select(litters_data, ends_with("weight"))
select(litters_data, contains("group, weight"))
```

## everything(), which is handy for reorganizing columns without discarding anything:

```{r}
select(litters_data, litter_number, pups_survive, everything())
```

## relocate does a similar thing:
```{r}
relocate(litters_data, litter_number, pups_survive)
```

## filter: "==":equal to, "!=": not equal to
##"%in%": see data is or not in the dataset,
```{r}
filter(litters_data, gd_of_birth == 20)
filter(litters_data, pups_survive >= 2)
filter(litters_data, group == "Con7" & gd_of_birth == 20)
filter(litters_data, group %in% c("Con7", "Con8"))
```

## drop_na: omit misssing obervations
```{r}
filter(drop_na(litters_data))
```

##mutate: change or create a new column
```{r}
#litter_data2 =
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
  # wt_gain_kg = wt_gain * 2.2
)
```


## arrange: 

```{r}
head(arrange(litters_data, desc(group), pups_born_alive), 10)
```

## %>%
```{r}
litters_data_raw = read_csv("data_import_examples/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = 
  mutate(
    litters_data_selected_cols, 
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))
litters_data_with_vars_without_missing = 
  drop_na(litters_data_with_vars, wt_gain)
litters_data_with_vars_without_missing
```

```{r}
litters_data = 
  read_csv("data_import_examples/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
   drop_na(wt_gain)
 litters_data
```
## %>% : ctrl+shift+m
## lm: linear model 
```{r}
litters_data %>% 
  lm(wt_gain ~ pups_born_alive, data =.) %>% 
  broom::tidy()
```

